Think deep for reef resilience

### 2025-09-12 paper/paper_draft.qmd
Adjust abstract formatting
Abstract also needs to be revisited
Rework reef definition paragraphs for flow
Simplify and focus narrative
Exploratory modelling is a feature of Robust Decision Making but probably muddles the narrative here.

Depth characteristics for each reef were extracted using 10m resolution bathymetry data made
available in GBRMPA [-@GBRMPA2020]. For each reef, bathymetry data were spatially matched
with the 3806 reef outlines through geometric intersections. All depth values were converted
to positive values (depths below sea level). The median depth was used to provide an
indicative depth value for each reef. Code, and the resulting depth estimates, are available
at [@iwanaga2025].

While the bathymetry data adopted for the given assessment are said to be accurate up to 15m
from Mean Sea Level [@GBRMPA2020], there are many sources of uncertainty in the depth
estimates used for this study. These are outlined further in the Discussion. Notably, the
adopted depth estimates were derived from bathymetries based on satellite imagery. The
bathymetry data can be limited in their alignment with the established GBRMPA reef outlines
used to characterise reef locations and their areal extent. A Quality Assessment and Control
process identified reefs with potential data limitations: reefs with no overlapping
bathymetry data, reefs with minimum depth values above sea level (indicating potential data
artifacts), and reefs where bathymetry pixels covered less than 5% of the reef polygon area.
This process removed 842 reefs, or 22%, from the initial set of 3806 reefs. Furthermore,
reefs that are not associated with GBRMPA Management areas [@GBRMPA2007] or reefal
bioregions [@GBRMPA2024] were also removed from the investigation. Bioregions which
contained fewer than three reefs after these processes were also excluded. As a result, 1377
reefs were excluded (36%), leaving a total of 2429 remaining for the purpose of this
investigation.

@@

We adapted exploratory scenario modelling approach from robust decision making (RDM)
approaches for this investigation. In particular, we used parameter sampling and exploratory
modelling to investigate a wide range of plausible futures under different assumptions of
environmental (climatic) conditions and ecological factors. Exploratory modelling approaches
have gained favour in recent years as precise long-term predictions of future conditions are
often impossible in complex ecological systems [@conradt2024; @lempert2013]. RDM embraces
uncertainty as an inherent characteristic of complex environmental systems [@lempert2010;
@lempert2013] and is a core methodology underpinning the development of robust and adaptive
management strategies [@haasnoot2013]. Rather than attempting to forecast specific outcomes,
RDM aims to systematically explore many possible futures to identify patterns. Allowing for
the identification of robust management strategies that perform well across those diverse
conditions, as well as conditions in which such strategies may fail. Multiple interacting
and compounding sources of uncertainty, and a lack of broad agreement as to their
likelihood, create inherently uncertain outcomes. While we did not directly model management
decision making in this investigation, we used exploratory modelling to identify the
environmental and reef characteristics, such as thermal stress, depth and connectivity, that
support favourable outcomes such as recovery and resilience.

>>>>

Depth characteristics for each reef were extracted using 10m resolution bathymetry data made
available in GBRMPA [-@GBRMPA2020]. While the bathymetry data adopted for the given
assessment are said to be accurate up to 15m from Mean Sea Level [@GBRMPA2020], there are
many sources of uncertainty in the depth estimates used for this study. Notably, the adopted
depth estimates were derived from bathymetries based on satellite imagery. The bathymetry
data can be limited in their alignment with the established GBRMPA reef outlines used to
characterise reef locations and their areal extent. For each reef, bathymetry data were
spatially matched with the 3806 reef outlines through geometric intersections.

A Quality Assessment and Control process identified reefs with potential data limitations:
reefs with no overlapping bathymetry data, reefs with minimum depth values above sea level
(indicating potential data artifacts), and reefs where bathymetry pixels covered less than
5% of the reef polygon area. This process removed 842 reefs, or 22%, from the initial set of
3806 reefs. Furthermore, reefs that are not associated with GBRMPA Management areas
[@GBRMPA2007] or reefal bioregions [@GBRMPA2024] were also removed from the investigation.
Bioregions which contained fewer than three reefs after these processes were also excluded.
As a result, 1377 reefs were excluded (36%), leaving a total of 2429 remaining for the
purpose of this investigation. All depth values were converted to positive values (depths
below sea level). The median depth was used to provide an indicative depth value for each
reef. Further limitations are outlined in the Discussion. Code to extract depth estimates,
and the resulting depth estimates themselves, are available at [@iwanaga2025].

@@

We adapted an exploratory scenario modelling approach for this investigation. Specifically,
we use combinations of model factors to investigate a wide range of potential futures under
different assumptions of environmental (climatic) and ecological conditions. Exploratory
modelling approaches have gained favour in recent years to explore the implications of deep
uncertainty; where there is little or no agreement on the key drivers of a system, the
extent of the relationship, how best to represent them or their probability distributions
[@conradt2024; @lempert2013]. Exploratory modelling embraces uncertainty as an inherent
characteristic of complex environmental systems [@lempert2010; @lempert2013] and is a core
methodology underpinning the development of robust and adaptive management strategies
[@haasnoot2013].

Rather than attempting to forecast precise outcomes, the exploratory approach aims to
systematically search many possible futures to identify patterns that lead to outcomes of
interest. The approach allows for the identification of robust management strategies that
perform well across a diversity of future conditions rather than attempting to identify a
single most optimal strategy. Concurrently, assessing the wide swathe of scenarios allows
contextual understanding of the conditions in which strategies may fail. Multiple
interacting and compounding sources of uncertainty, and a lack of broad agreement as to
their likelihood, create inherently uncertain outcomes. Here, exploratory modelling was
applied to identify the environmental and reef characteristics, such as thermal stress,
depth, and connectivity, that support favourable outcomes in reef recovery and resilience.
